---
title: "Jordan-Analysis"
output: html_document
date: "2025-08-03"
editor_options: 
  chunk_output_type: inline
---

```{r load packages, echo=FALSE}
library(ggplot2)
library(dplyr)
library(data.table)
library(lmtest)
library(sandwich)
library(stargazer)
```

## Distribution of outcome variable

```{r outcome distribution}
d_subset <- read.csv('data/data.csv')

hist(d_subset$purchase_amount,
     breaks = c(20),
     main = 'Distribution of Purchase Amounts',
     xlab = 'Purchase Amount',
     col = 'darkcyan')
```

## Statistical Analysis

### Group means

```{r group means}
group_means <- d_subset[ , .(mean_purchase_amt = mean(purchase_amount)), 
                  keyby = PresentationType]

group_means
```

*Note*: Just observing the averages for each treatment assignment group, it looks like the Generic_Individual group actually has the highest average purchase amount, with Accurate_Bundle having the lowest average purchase amount. 

### Regressions of group means for purchase amount

Regression of treatment group means

```{r regression of means}
# relevel groups to set baseline comparison
d_subset$PresentationType <- as.factor(d_subset$PresentationType)

d_subset$PresentationType <- relevel(d_subset$PresentationType, ref = 'Generic_Individual')

model_means <- lm(purchase_amount ~ PresentationType, data = d_subset)

stargazer(model_means, type = 'text', 
          title = 'Results: Purchase Amount by Treatment Group',
          dep.var.labels = 'Purchase Amount',
          covariate.labels = c('Accurate Bundle', 'Accurate Individual', 'Generic Bundle',
                               'Inaccurate Bundle', 'Inaccurate Individual', 
                               'Constant (Generic Individual)'),
          ci = TRUE,
          font.size = 'large',
          no.space = TRUE)
```


*Note*: None of the results appear to be statistically significant. The regression does confirm that those in the Generic_Individual group (the intercept reference group) spent more than other groups. Those in the Generic_Bundle group (non-personalized bundle) spent about `r round(model_simple$coefficients[2], 2)` less than the baseline, and those in the Accurate_Bundle (personalized bundle) group spent about `r round(model_simple$coefficients[4], 2)` less than the baseline. 

#### With demographics

```{r demographic regression of means}
model_demographics <- lm(purchase_amount ~ PresentationType + Age + Gender, data = d_subset)

stargazer(model_means, model_demographics,
          type = 'text',
          title = 'Results: Purchase Amount by Treatment Group with Demographics',
          dep.var.labels = 'Purchase Amount',
          covariate.labels = c('Accurate Bundle', 'Accurate Individual', 'Generic Bundle',
                               'Inaccurate Bundle', 'Inaccurate Individual'),
          ci = TRUE,
          font.size = 'large',
          no.space = TRUE,
          omit = c("Age", "Gender"),
          notes = c("Model 2 represents controls for age and gender.")
)
```


### Regression of bundling and personalization interaction for purchase amount

```{r interaction regression}
d_subset <- d_subset %>% #create Personalization_Treatment column
  mutate(Personalization_Treatment = sub("_.*", "", PresentationType))

d_subset$bundle <- factor(d_subset$bundle, levels = c(0, 1))  # 0 = reference
d_subset$Personalization_Treatment <- relevel(factor(d_subset$Personalization_Treatment), ref = "Generic")

model_full <- lm(purchase_amount ~ bundle*Personalization_Treatment, data = d_subset)
robust_se1 <- sqrt(diag(vcovHC(model_full, type = 'HC1')))

stargazer(model_full, type = 'text', 
          title = 'Results: Effects of Bundling and Personalization',
          dep.var.labels = 'Purchase Amount',
          se = list(robust_se1),
          ci = TRUE,
          font.size = 'large',
          no.space = TRUE)
```

#### With demographics

```{r interaction regression with demographics}
model_full_demographics <- lm(purchase_amount ~ bundle*Personalization_Treatment + Age + Gender + Q15, data = d_subset)
robust_se2 <- sqrt(diag(vcovHC(model_full_demographics, type = "HC1")))

stargazer(model_full, model_full_demographics,
          type = 'text', 
          title = 'Results: Effects of Bundling and Personalization',
          dep.var.labels = 'Purchase Amount',
          se = list(robust_se2),
          ci = TRUE,
          font.size = 'large',
          no.space = TRUE,
          omit = c("Age", "Gender", "Q15"),
          notes = c("Model 2 represents controls for age and gender."))
```

```{r}
anova(model_full, model_full_demographics)
```

## All regressions for purchase amount

```{r full regression}
######### Prepare dataset ######### 
d_subset <- d_subset %>% #create Personalization_Treatment column
  mutate(Personalization_Treatment = sub("_.*", "", PresentationType))

d_subset$bundle <- factor(d_subset$bundle, levels = c(0, 1))  # 0 = reference
d_subset$Personalization_Treatment <- relevel(factor(d_subset$Personalization_Treatment), ref = "Generic")

######### bundling model ######### 
model_bundling <- d_subset[, lm(purchase_amount ~ bundle)]
bundling_rse <- sqrt(diag(vcovHC(model_bundling, type="HC1")))

##########  personalization model ######### 
model_personalization <- d_subset[, lm(purchase_amount ~ Personalization_Treatment)]
personalization_rse <- sqrt(diag(vcovHC(model_personalization, type="HC1")))

##########  Interaction model ######### 
model_full <- lm(purchase_amount ~ bundle*Personalization_Treatment, data = d_subset)
full_rse <- sqrt(diag(vcovHC(model_full, type = 'HC1')))

##########  Interaction with demographics ######### 
model_full_demographics <- lm(purchase_amount ~ bundle*Personalization_Treatment + Age + Gender + Q15, data = d_subset)
demographics_rse <- sqrt(diag(vcovHC(model_full_demographics, type = "HC1")))


######### Stargazer ######### 
stargazer(model_bundling, model_personalization, model_full, model_full_demographics,
          se = list(bundling_rse, personalization_rse, full_rse, demographics_rse),
          type = 'text', 
          title = 'Results: Effects of Bundling and Personalization',
          dep.var.labels = 'Purchase Amount',
          ci = TRUE,
          font.size = 'large',
          no.space = TRUE,
          omit = c("Age", "Gender", "Q15"),
          notes = c("Model 4 represents same interaction as Model 3 between bundling and personalization with the addition of controls for age, gender and income."))
```

```{r}
anova(model_bundling, model_personalization, model_full, model_full_demographics)

```

## Regressions for binary purchase outcome (yes= 1, no= 0)

```{r}
# Binary outcome column
d_subset$if_purchase <- ifelse(d_subset$purchase_amount > 0, 1, 0)

# Models
model_bundle_y2 <- glm(if_purchase ~ bundle, data = d_subset, family = binomial(link = "logit"))

model_personalization_y2 <- glm(if_purchase ~ Personalization_Treatment, data = d_subset, family = binomial(link = "logit"))

model_full_y2 <- glm(if_purchase ~ Personalization_Treatment * bundle, data = d_subset, family = binomial(link = "logit"))

model_full_demographics_y2 <- glm(if_purchase ~ Personalization_Treatment * bundle + Age + Gender + Q15, data = d_subset, family = binomial(link = "logit"))

# Function to get robust SE safely
robust_se <- function(model) {
  sqrt(diag(vcovHC(model, type = "HC1")))
}

# Fit models (after cleaning data)
model_list <- list(
  model_bundle_y2,
  model_personalization_y2,
  model_full_y2,
  model_full_demographic_y2
)

# Calculate SEs
se_list <- lapply(model_list, robust_se)

# Output
stargazer(model_list, se = se_list,
          type = "text",
          title = "Logistic Regression Results with Robust SE",
          dep.var.labels = "Any purchase (1=yes, 0=no)",
          digits = 3, 
          header = FALSE,
          ci = TRUE,
          font.size = 'large',
          no.space = TRUE,
          omit = c("Age", "Gender", "Q15"),
          notes = c("Model 4 represents same interaction as Model 3 between bundling and personalization with the addition of controls for age, gender and income."))

```

## Full table purchase and binary outcome

```{r}
###### Define models #####
outcome_y1_models <- list(model_bundling, model_personalization,
                          model_full, model_full_demographics)
outcome_y2_models <- list(model_bundle_y2, model_personalization_y2,
                          model_full_y2, model_full_demographic_y2)

all_models <- c(outcome_y1_models, outcome_y2_models)

##### Robust SEs #####
se_list_all <- lapply(all_models, robust_se) # apply function from above

##### Stargazer table #####
stargazer(all_models,
          se = se_list_all,
          type = 'text',
          title = "Effects on Purchase Amount and Likelihood of Purchase",
          column.labels = c('Purchase Amount', 'Any Purchase (logit)'),
          column.separate = c(length(outcome_y1_models), length(outcome_y2_models)),
          model.numbers = FALSE,
          digits = 2,
          ci = TRUE,
          ci.level = 0.95,
          omit = c('Age', 'Gender', 'Q15'),
          notes = c(
            "Robust (HC1) standard errors in parentheses; 95% confidence intervals shown.", "Columns 1–4: OLS on purchase amount (in currency units). Columns 5–8: Logistic regression on any purchase (1=yes, 0=no)."
  ),
  notes.align = "l")
```


## Appendix

Non-causal interaction effect of perceived personalization and bundle on purchase amount.

```{r simple model using indicator variables}
model_encoded <- lm(purchase_amount ~ bundle*personalization, data = d_subset)
summary(model_encoded)
```


```{r encoded model and demographics}
# Non-causal interaction effect of perceived personalization and bundle on purchase amount
model_encoded_demographics <- lm(purchase_amount ~ bundle*personalization + Age + Gender, data = d_subset)

stargazer(model_encoded, model_encoded_demographics,
          type = 'text',
          title = 'Bundle and Personalization interaction with demographics',
          dep.var.labels = 'Purchase Amount',
          ci = TRUE,
          font.size = 'large',
          no.space = TRUE,
          omit = c("Age", "Gender"),
          notes = c("Model 2 represents controls for age and gender."))

```

